---
title: "PhotoCold"
author: "Yunpeng Luo"
date: "5/28/2021"
output: html_document
---
## Description

**After exploring the general pattern of modelling GPP vs observational GPP, the next step to identify the specific period when the mismatch between modeled GPP and observed GPP in each site-->focused in the markdown file**

step1: tidy the table for GPP simulation vs GPP obs sites

step2: finding the way to separate out the model early simulation period

## step1: tidy the table
```{r}
library(kableExtra)
library("readxl")
table.path<-"C:/Users/yluo/Desktop/CES/Data_for_use/"
my_data <- read_excel(paste0(table.path,"Info_Table_about_Photocold_project.xlsx"), sheet = "Only_sites_earlyGPPest")
# my_data %>%
# kbl(caption = "Summary of sites with early GPP estimation") %>%
#   kable_paper(full_width = F, html_font = "Cambria") %>%
#   scroll_box(width = "500px", height = "200px") #with a scroll bars
my_data %>%
  kbl(caption = "Summary of sites with early GPP estimation") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```


## step2: seprate the time period when model early estimation of GPP

### Part1: find the method to determine the period that with early GPP estimation

```{r echo = FALSE, message = FALSE, warning = FALSE,fig.height=10,fig.width=10}
library(readr)
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(cowplot)

path<-"C:/Users/yluo/Desktop/CES/Data_for_use/Data_sent_by_Beni/"
setwd(path)
ddf <- read_csv(paste0(path,"/ddf_fluxnet2015_pmodel_with_forcings_stocker19gmd.csv"))
#convert the format of date
#ddf$date<-strptime(ddf$date,format = "%d/%m/%Y")
ddf$date<-as.Date(ddf$date,"%m/%d/%Y")

# goodsites <- c("IT-Ren", "RU-Ha1", "US-Me2", "IT-SRo", "BE-Vie", "CH-Cha", "CH-Lae", "CH-Oe1", "DE-Gri", "DE-Obe", "DE-RuR", "DE-Tha", "NL-Hor", "NL-Loo")
#as the US_Me2 and IT_SRo belongs to Csb and Csa, for the analysis, only keep Cfb,Dfb, and Cfc.
#Hence,remove these two sites for analysis 
goodsites <- c("IT-Ren", "RU-Ha1", "BE-Vie", "CH-Cha", "CH-Lae", "CH-Oe1", "DE-Gri", "DE-Obe", "DE-RuR", "DE-Tha", "NL-Hor", "NL-Loo")
badsites <- c("US-Syv", "US-UMB", "US-UMd", "US-WCr", "US-Wi3", "CA-Man", "CA-NS2", "CA-NS4", "CA-NS5", "CA-Qfo", "FI-Hyy", "IT-Tor", "DE-Hai") 
####
library(phenopix)
library(zoo)
#function: to smooth the time series
extract_smooth_y<-function(x,df){
  # x<-df_new$gpp_mod_FULL_f
  # df<-0.05
  x_sm<-SplineFit(x,df=df)
  y<-x_sm$fit$predicted
  return(y)
}
#function: to determine the green-up period
#function based on the threshold:
Deter_greenup_thrs<-function(df,pos_max){
  # df<-df_subset
  # pos_max<-pos_max

  df_sel_sos<-df %>%
    filter(doy<=pos_max)
  #1) calculate the amplitude of the time series
  #based on the 10% or 25% of amplitude:
  mn_sos<-min(df_sel_sos$gpp_mod_norm_sm,na.rm = T)
  mm<-max(df_sel_sos$gpp_mod_norm_sm,na.rm = T)
  amp_sos<-mm-mn_sos
  
  trs_10_sos<-mn_sos+0.1*amp_sos
  trs_25_sos<-mn_sos+0.25*amp_sos
   
  #2)determine the start of season(sos10, and sos25)
  sos10<-which.min(abs(as.numeric(df_sel_sos$gpp_mod_norm_sm - trs_10_sos)))
  sos25<-which.min(abs(as.numeric(df_sel_sos$gpp_mod_norm_sm - trs_25_sos)))
  #3)the end of the green-up period
  peak<-pos_max
  #4)additional:also determine eos10 and eos25
  df_sel_eos<-df %>%
    filter(doy>=pos_max)
  mn_eos<-min(df_sel_eos$gpp_mod_norm_sm,na.rm = T)
  amp_eos<-mm-mn_eos
  #
  trs_10_eos<-mn_eos+0.1*amp_eos
  trs_25_eos<-mn_eos+0.25*amp_eos
  #
  eos10<-which.min(abs(as.numeric(df_sel_eos$gpp_mod_norm_sm - trs_10_eos)))
  eos25<-which.min(abs(as.numeric(df_sel_eos$gpp_mod_norm_sm - trs_25_eos)))
  #adjust the eos 10 and eos25
  eos25<-eos25+peak-1
  eos10<-eos10+peak-1
  #
  pos_agg<-data.frame(sos10=sos10,sos25=sos25,peak=peak,eos25=eos25,eos10=eos10)
  #plotting demonstration:
  # plot(df_subset$date,df$gpp_mod_norm,ylab = "Norm GPP",xlab = "",pch=".")
  # lines(df_subset$date,df$gpp_mod_roll20,ylab = "Norm GPP",xlab = "",lwd=1.2,col="red")
  # abline(h=c(trs_10_sos,trs_25_sos,trs_25_eos,trs_10_eos),col="red",lty=2)
  # abline(v=df_subset$date[c(sos10,sos25,eos25,eos10)],col="red")
  # text(df_subset$date[c(sos10,sos25,eos25,eos10)],0.8,labels = c("sos10","sos25","eos25","eos10"),col = "red",pos = c(2,4,2,4))
  # abline(v=df_subset$date[c(peak)],col="red")
  # text(df_subset$date[c(peak)],0.8,labels = c("peak"),col = "red",pos = c(4))
  return(pos_agg)
}

#here using the klosterman method to extract the phenopohases 
Deter_greenup<-function(df,pos_max){
  # df<-df_subset
  # pos_max<-pos_max

  #1)extract the phenophase from time series(only using the GPPmod)
  #using the Klosterman et al., 2014 to fit the time series
  #followed the function in Phenopix R package  
  
  #convert to zoo
  df_sel<-df
  gpp_mod_norm_sm<-zoo(df_sel$gpp_mod_norm_sm,order.by = as.POSIXct(strptime(df_sel$date,format = "%Y-%m-%d")))
  
  #fit the data 
  gpp_mod_fit<-KlostermanFit(gpp_mod_norm_sm,which="light")
  
  #extract the phenophases:
  gpp_mod_phenos<-PhenoKl(x=gpp_mod_fit$fit$params,
                          fit=gpp_mod_fit$fit)

  #2)merge the phenos
  peak<-pos_max
  pos_agg<-data.frame(sos=gpp_mod_phenos[1],Maturity=gpp_mod_phenos[2],peak=peak,
                      Senescence=gpp_mod_phenos[3],eos=gpp_mod_phenos[4],method="kls")
  #3)adjustment-->if klosterman method does not work, using thrs method
  if(length(pos_agg[!is.na(pos_agg)])<5){
    pos_agg<-Deter_greenup_thrs(df_sel,pos_max)
    pos_agg<-data.frame(pos_agg,method="thrs")
    names(pos_agg)<-c("sos","Maturity","peak","Senescence","eos","method")
    # print("using thrs")
  }
  #plotting demonstration:
  # plot(df_subset$date,df$gpp_mod_norm,ylab = "Norm GPP",xlab = "",pch=".")
  # lines(df_subset$date,df$gpp_mod_roll20,ylab = "Norm GPP",xlab = "",lwd=1.2,col="red")
  # abline(v=df_subset$date[unlist(pos_agg)],col="red")
  # text(df_subset$date[unlist(pos_agg)],0.8,labels = c("sos","maturity","peak","senescence","eos"),col = "red",pos = c(2,2,4,4,4))
  return(pos_agg)
}

##function to seprate the early GPP estimation period:
sep_data_indiffY_sepMismatch<-function(ddf,degf,Clim,PFT,site_name,avai_years){
    # ddf<-ddf
    # degf<-0.05  #degree of freedom for the spline smooth
    # Clim<-"Cfb"
    # PFT<-"DBF"
    # site_name<-"DE-Hai"
    # avai_years<-c(2000:2012)
    ##
    df<-ddf %>%
    filter(sitename==site_name) %>%
    mutate(doy = yday(date), Year=year(date)) %>%
    filter(Year %in% avai_years)
    #
    Years_unique<-unique(df$Year)
  
    #-------------------------------------------------------------------------------
    #Step1: normlization for all the years in one site
    #normalized the gpp_obs and gpp_mod using the gpp_max(95 percentile of gpp)
    #-------------------------------------------------------------------------------
    gpp_max_obs<-quantile(df$gpp_obs,0.95,na.rm = T)
    gpp_max_mod<-quantile(df$gpp_mod_FULL,0.95,na.rm = T)
    df$gpp_mod_norm<-df$gpp_mod_FULL/gpp_max_mod
    df$gpp_obs_norm<-df$gpp_obs/gpp_max_obs
    #first to gap-fill the ts to enable the ts smoothing
    #then smooth the time series of all the years uisng the spline(did not use at the moment)
    #hints:do not gap-fill gpp_obs if there are long gaps in the time sereis-->for biases calculation in step 4-->need to gapfill each year
    if(length(df$gpp_obs[!is.na(df$gpp_obs)])>30){
      df_new<-c()
      for(i in 1:length(Years_unique)){
      df_t<-df[df$Year==Years_unique[i],]
      df_t<-df_t %>%
          mutate(gpp_mod_norm_f=na.fill(gpp_mod_norm,c("extend")),gpp_obs_norm_f= na.fill(gpp_obs_norm,c(NA,"extend",NA))) %>%
      mutate(gpp_mod_norm_sm = extract_smooth_y(gpp_mod_norm_f,degf))
      df_new<-rbind(df_new,df_t)
      }
    }
    #
    N<-length(df$gpp_mod_FULL)
    #-------------------------------------------------------------------------------
    #Step 2:Determine the green-up period for each year(using spline smoothed values): 
    #followed analysis is based on the normlized "GPP_mod"time series(determine earlier sos)
    #-------------------------------------------------------------------------------
    #
    library(zoo)
    pos_diffYears<-c()
    #also subset data beyond the green-up period
    df_outgreenup<-c()
    df_final<-c()
    for(i in 1:length(Years_unique)){
    df_subset<-df_new[df_new$Year==Years_unique[i],]
    #calculate the rollmean 
    df_subset<-df_subset%>%
        mutate(gpp_mod_roll5=rollapply(df_subset$gpp_mod_norm_f,5,mean,fill=NA,align="center"),
               gpp_mod_roll7=rollapply(df_subset$gpp_mod_norm_f,7,mean,fill=NA,align="center"),
               gpp_mod_roll10=rollapply(df_subset$gpp_mod_norm_f,10,mean,fill=NA,align="center"),
               gpp_mod_roll15=rollapply(df_subset$gpp_mod_norm_f,15,mean,fill=NA,align="center"),
               gpp_mod_roll20=rollapply(df_subset$gpp_mod_norm_f,20,mean,fill=NA,align="center"),
               gpp_obs_roll5=rollapply(df_subset$gpp_obs_norm_f,5,mean,fill=c(NA,"extend",NA),align="center"),
               gpp_obs_roll7=rollapply(df_subset$gpp_obs_norm_f,7,mean,fill=c(NA,"extend",NA),align="center"),
               gpp_obs_roll10=rollapply(df_subset$gpp_obs_norm_f,10,mean,fill=c(NA,"extend",NA),align="center"),
               gpp_obs_roll15=rollapply(df_subset$gpp_obs_norm_f,15,mean,fill=c(NA,"extend",NA),align="center"),
               gpp_obs_roll20=rollapply(df_subset$gpp_obs_norm_f,20,mean,fill=c(NA,"extend",NA),align="center"),
               temp_obs_roll20=rollapply(df_subset$temp_day_fluxnet2015,20,mean,fill=c(NA,"extend",NA),align="center")
               )
    # pos_sim_max<-match(max(df_subset$gpp_mod_roll20,na.rm = T),df_subset$gpp_mod_roll20)
    pos_sim_max<-match(max(df_subset$gpp_mod_norm_sm),df_subset$gpp_mod_norm_sm)
    pos_max<-pos_sim_max
    #finding out the start and end of the green-up period:
    pos_agg<-Deter_greenup(df_subset,pos_max)
    pos_diffYears<-rbind(pos_diffYears,pos_agg)
    #separate the data beyond greenup period:
    # pos_beyond<-c(1:c(pos_agg$sos10-1),c(c(pos_agg$peak+1):length(df_subset$date)))
    #pos_beyond<-c(pos_agg$peak+1):length(df_subset$date)
    pos_beyond<-c(1:c(pos_agg$sos-1),c(c(pos_agg$eos+1):length(df_subset$date)))
    t_outgreenup<-df_subset[pos_beyond,]
    df_outgreenup<-rbind(df_outgreenup,t_outgreenup)
    df_final<-rbind(df_final,df_subset)
    }
    rownames(pos_diffYears)<-Years_unique
    #-------------------------------------------------------------------------------
    #Step 3:rolling mean of GPPobs and GPPmod for data for all the years(moving windown:5,7,10, 15, 20days)
    #also for the data beyond green-up period-->moving to second step
    #-------------------------------------------------------------------------------
    # df_outgreenup<-df_outgreenup%>%
    #     mutate(gpp_mod_roll5=rollapply(df_outgreenup$gpp_mod_norm,5,mean,na.rm=T,fill=NA,align="center"),
    #            gpp_mod_roll7=rollapply(df_outgreenup$gpp_mod_norm,7,mean,na.rm=T,fill=NA,align="center"),
    #            gpp_mod_roll10=rollapply(df_outgreenup$gpp_mod_norm,10,mean,na.rm=T,fill=NA,align="center"),
    #            gpp_mod_roll15=rollapply(df_outgreenup$gpp_mod_norm,15,mean,na.rm=T,fill=NA,align="center"),
    #            gpp_mod_roll20=rollapply(df_outgreenup$gpp_mod_norm,20,mean,na.rm=T,fill=NA,align="center"),
    #            gpp_obs_roll5=rollapply(df_outgreenup$gpp_obs_norm,5,mean,na.rm=T,fill=NA,align="center"),
    #            gpp_obs_roll7=rollapply(df_outgreenup$gpp_obs_norm,7,mean,na.rm=T,fill=NA,align="center"),
    #            gpp_obs_roll10=rollapply(df_outgreenup$gpp_obs_norm,10,mean,na.rm=T,fill=NA,align="center"),
    #            gpp_obs_roll15=rollapply(df_outgreenup$gpp_obs_norm,15,mean,na.rm=T,fill=NA,align="center"),
    #            gpp_obs_roll20=rollapply(df_outgreenup$gpp_obs_norm,20,mean,na.rm=T,fill=NA,align="center")
    #            )
    #-------------------------------------------------------------------------------
    #Step 4:Fit the Guassian norm distribution for residuals beyond the green-up period
    #-------------------------------------------------------------------------------
    library(MASS)
    #calculate the residuals
    res.gpp<-data.frame(date=df_outgreenup$date,res=c(df_outgreenup$gpp_mod_roll20 - df_outgreenup$gpp_obs_roll20))
    res.gpp<-res.gpp[!is.na(res.gpp$res),]
    fit_res<-fitdistr(res.gpp$res,"normal")
    para <- fit_res$estimate
    # hist(res_gpp,probability = TRUE)
    # curve(dnorm(x, para[1], para[2]), col = 2, add = TRUE)
    mu<-mean(res.gpp$res,na.rm = T)
    p_density_res_outgreenup<-ggplot(res.gpp, aes(x=res))+
      geom_histogram(aes(y=..density..),binwidth = 0.01,position="identity", alpha=0.3,fill="white",col="green3",)+
      geom_density(col="forestgreen",size=1.05)+
      stat_function(fun = dnorm, args = list(mean = mean(res.gpp$res), sd = sd(res.gpp$res)),size=1.05,col="black")+
      geom_vline(xintercept=mu,linetype="dashed",size=1.05)+
      theme_minimal()
    #-------------------------------------------------------------------------------
    #step 5:determine the "is_event" within green-up period
    #-------------------------------------------------------------------------------
    df_diffYears<-c()
    for(i in 1:length(Years_unique)){
    df_subset<-df_final[df_final$Year==Years_unique[i],]
    #a.add one column to indicate the "greenup" 
    df_subset$greenup<-rep("no",length(df_subset$date))
    pos_agg<-pos_diffYears[i,]
    #green-up date=yes
    df_subset[c(df_subset$doy>=pos_agg$sos) & c(df_subset$doy<=pos_agg$peak),]$greenup<-"yes"
    
    #b.add one column to indciate "is_event"
    #set the biases(gppmod-gppobs) are positive and bigger than 1.2sd as the "is_event"
    df_subset$gpp_res<-df_subset$gpp_mod_roll20-df_subset$gpp_obs_roll20
    #remove the NA
    # df_subset_rmNA<-df_subset[!is.na(df_subset$gpp_res),]
    thrs<-para[2]*1.2
    #"is_event"should be in the green-up period
    df_subset$is_event<-rep("no",length(df_subset$date))
    #for the data is not equal to NA
    df_subset[c(df_subset$greenup=="yes")&c(df_subset$gpp_res>thrs)&c(!is.na(df_subset$gpp_res)),]$is_event<-"yes"
    #for the data is equal to NA
    df_subset[c(df_subset$greenup=="yes")&c(df_subset$gpp_res>thrs)&c(is.na(df_subset$gpp_res)),]$is_event<-"NA"
    ############################
    #adjusted the "is_event"
    pos_isevent<-df_subset$doy[which(df_subset$is_event=="yes")]
    ##plotting
    # par(mfrow=c(2,1))
    # plot(df_subset$date,df_subset$temp_day_fluxnet2015,pch=1)
    # points(df_subset$date,df_subset$temp_obs_roll20,col="red")
    # abline(h=c(0,10,15),col="red")
    # #
    # plot(df_subset$date,df_subset$gpp_mod_roll20,col="red",ylim=c(-0.1,1.2))
    # points(df_subset$date,df_subset$gpp_obs_roll20)
    # abline(v=df_subset$date[c(min(pos_isevent),max(pos_isevent))])

    #method: find the postion of diff_pos_isevent>10;remove the "pos_isevent" after the positon (do not use this at the moment)
    # diff_pos_isevent<-diff(pos_isevent)
    # t_pos<-match(diff_pos_isevent[diff_pos_isevent>10],diff_pos_isevent)
    # if(length(t_pos)>0){
    # rm_pos_isevent<-pos_isevent[t_pos+1]
    # #adjust the pos_isevent:
    # pos_isevent_new<-pos_isevent[pos_isevent<rm_pos_isevent]
    # ##set flag of other position beyond "is_event" as "no"
    # pos_rm<-setdiff(pos_isevent,pos_isevent_new)
    # df_subset$is_event[pos_rm]<-rep("no",length(pos_rm))
    # }
    #method:remove the positon of pos_isevent when air temperature >10
    pos_temp_less10<-df_subset$doy[which(c(df_subset$is_event=="yes") & c(df_subset$temp_obs_roll20<=10))]
    # abline(v=df_subset$date[c(min(pos_temp_less10),max(pos_temp_less10))],col="red")
    df_subset$is_event_less10<-df_subset$is_event
    pos_rm<-setdiff(pos_isevent,pos_temp_less10)
    df_subset$is_event_less10[pos_rm]<-rep("no",length(pos_rm))
    ##############################
    df_subset$date<-as.POSIXct(df_subset$date)
    #c.merge the data.frame
    df_diffYears<-rbind(df_diffYears,df_subset)
    }
    #-------------------------------------------------------------------------------
    #step 6:Evaluation "is_event"-->visualization and stats
    #-------------------------------------------------------------------------------
    #plotting+stats
    p_diffYears<-c()
    Pfalse_diffYears<-c()
    for(i in 1:length(Years_unique)){
    df_subset<-df_diffYears[df_diffYears$Year==Years_unique[i],]
    #a.stats:positive false%= days(gpp_mod underestimation for days flaged as is_event)/days(is_event)
    df_stats<-subset(df_subset,is_event=="yes"&!is.na(gpp_obs))
    df_stats$flag<-rep("overEsti",length(df_stats$date))
    for(k in 1:length(df_stats$date)){
      ifelse(df_stats$gpp_mod_FULL[k]-df_stats$gpp_obs[k]<0,df_stats$flag[k]<-"underEsti",df_stats$flag[k]<-df_stats$flag[k])
    }
    Pfalse<-round(c(length(which(df_stats$flag=="underEsti"))/length(df_stats$date)*100),1)
    Pfalse_diffYears<-rbind(Pfalse_diffYears,Pfalse)
    #b.plotting:demonstration the original value + moving average + dividing lines
    df_subset$date<-as.POSIXct(df_subset$date)
    pos_agg<-pos_diffYears[i,]
    pos_isevent<-df_subset$doy[which(df_subset$is_event=="yes")]
    pos_isevent_less10<-df_subset$doy[which(df_subset$is_event_less10=="yes")]
    rect.coord_greenup=data.frame(x1=df_subset$date[pos_agg[,"sos"]],x2=df_subset$date[pos_agg[,"peak"]], y1=min(df_subset$gpp_mod_norm,na.rm = T)-0.05, y2=max(df_subset$gpp_mod_norm,na.rm = T)+0.02)
    # rect.coord_isevent=data.frame(x1=df_subset$date[min(pos_isevent)],x2=df_subset$date[max(pos_isevent)], y1=min(df_subset$gpp_mod_norm,na.rm = T)-0.05, y2=max(df_subset$gpp_mod_norm,na.rm = T)+0.02)
    rect.coord_isevent_less10=data.frame(x1=df_subset$date[min(pos_isevent_less10)],x2=df_subset$date[max(pos_isevent_less10)], y1=min(df_subset$gpp_mod_norm,na.rm = T)-0.05, y2=max(df_subset$gpp_mod_norm,na.rm = T)+0.02)
#
  
  gg_plot<-ggplot() + 
      geom_point(data=df_subset,aes(x=date,y = gpp_mod_norm),col=adjustcolor("tomato",0.5),pch=1)+
      geom_line(data=df_subset,aes(x=date,y = gpp_mod_roll20,col="modelled"),lwd=1.2)+
      geom_point(data=df_subset,aes(x=date,y = gpp_obs_norm),col=adjustcolor("black",0.5),pch=1)+
      geom_line(data=df_subset,aes(x=date,y = gpp_obs_roll20,col="observed"),lwd=1.2)+
      geom_vline(xintercept = df_subset$date[pos_agg[,"sos"]],col="forestgreen",lty=2,lwd=1.2)+
      geom_vline(xintercept = df_subset$date[pos_agg[,"peak"]],col="forestgreen",lty=2,lwd=1.2)+
      scale_color_manual("",values = c("modelled"="red","observed"="black")) +
      labs(title = paste0(Clim,"-",PFT,"-",site_name,":",Years_unique[i]), x = "", y = "norm GPP" )+  #(g C m-2 d-1 )
  geom_rect(data=rect.coord_greenup,mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill="green3",alpha=0.1)+
  geom_vline(xintercept = df_subset[df_subset$is_event=="yes",]$date,col=adjustcolor("steelblue2",0.3))+
  # geom_rect(data=rect.coord_isevent,mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill="steelblue2",alpha=0.3)+
  geom_rect(data=rect.coord_isevent_less10,mapping=aes(xmin=x1, xmax=x2, ymin=y1, ymax=y2), fill="orange",alpha=0.5)+
  theme_classic()+
  theme(legend.position = c(0.15,0.9),legend.background = element_blank())
  
  ## merge the plot
  ifelse(i==1,gg_plot<-gg_plot+theme(),gg_plot<-gg_plot+theme(legend.position = "none"))
  # plot(gg_plot)
  p_diffYears[[i]]<-gg_plot
    }
  Pfalse_diffYears<-as.data.frame(Pfalse_diffYears)
  names(Pfalse_diffYears)<-"Pflase";rownames(Pfalse_diffYears)<-Years_unique
    #-------------------------------------------------------------------------------
    #step 7:Summary
    #-------------------------------------------------------------------------------
  df_agg<-c()
  df_agg$pos_agg<-pos_diffYears
  df_agg$df_agg<-df_diffYears
  df_agg$p_den_outgreenup<-p_density_res_outgreenup
  df_agg$Pflase<-Pfalse_diffYears
  df_agg$p_isevent<-p_diffYears
  return(df_agg)
}

##Cfb-DE-Hai
df_andPlot_Cfb<-sep_data_indiffY_sepMismatch(ddf,0.05,"Cfb","DBF","DE-Hai",c(2000:2012))
length(df_andPlot_Cfb$p_isevent)
p_diffYears<-df_andPlot_Cfb$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],p_diffYears[[3]],
          p_diffYears[[4]],p_diffYears[[5]],p_diffYears[[6]],
          p_diffYears[[7]],p_diffYears[[8]],p_diffYears[[9]],
          p_diffYears[[10]],p_diffYears[[11]],p_diffYears[[12]],
          p_diffYears[[13]],
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)
```

### Part2: check all the sites 
**(1) For Dfb:both for MF and DBF sites**
- Dfb-MF (1 site)
```{r echo = FALSE, message = FALSE, warning = FALSE,fig.height=10,fig.width=10}
#US-Syv
df_andPlot_Dfb_MF<-sep_data_indiffY_sepMismatch(ddf,0.05,"Cfb","DBF","US-Syv",c(2002:2006,2014))
length(df_andPlot_Dfb_MF$p_isevent)
p_diffYears<-df_andPlot_Dfb_MF$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],p_diffYears[[3]],
          p_diffYears[[4]],p_diffYears[[5]],p_diffYears[[6]],
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)
```

- Dfb-DBF (4 sites)
```{r echo = FALSE, message = FALSE, warning = FALSE,fig.height=10,fig.width=10}
#US-UMB
df_andPlot_Dfb_DBF1<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfb","DBF","US-UMB",c(2000:2014))
length(df_andPlot_Dfb_DBF1$p_isevent)
p_diffYears<-df_andPlot_Dfb_DBF1$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],p_diffYears[[3]],
          p_diffYears[[4]],p_diffYears[[5]],p_diffYears[[6]],
          p_diffYears[[7]],p_diffYears[[8]],p_diffYears[[9]],
          p_diffYears[[10]],p_diffYears[[11]],p_diffYears[[12]],
          p_diffYears[[13]],p_diffYears[[14]],p_diffYears[[15]],        
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)

#US-UMd
df_andPlot_Dfb_DBF2<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfb","DBF","US-UMd",c(2008:2014))
length(df_andPlot_Dfb_DBF2$p_isevent)
p_diffYears<-df_andPlot_Dfb_DBF2$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],p_diffYears[[3]],
          p_diffYears[[4]],p_diffYears[[5]],p_diffYears[[6]],
          p_diffYears[[7]],      
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)

#US-Wcr
df_andPlot_Dfb_DBF3<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfb","DBF","US-WCr",c(2000:2006,2011:2014))
length(df_andPlot_Dfb_DBF3$p_isevent)
p_diffYears<-df_andPlot_Dfb_DBF3$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],p_diffYears[[3]],
          p_diffYears[[4]],p_diffYears[[5]],p_diffYears[[6]],
          p_diffYears[[7]],p_diffYears[[8]],p_diffYears[[9]],
          p_diffYears[[10]],p_diffYears[[11]],
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)

#US-Wi3-->lack early spring observation
# df_andPlot_Dfb_DBF4<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfb","DBF","US-WCr",c(2002,2004))
# length(df_andPlot_Dfb_DBF4$p_agg)
# p_diffYears<-df_andPlot_Dfb_DBF4$p_agg
# p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],
#           labels = "auto",ncol=3,label_size = 12,align = "hv")
# plot(p_merge)
```

**(2) For Dfc:both for GRA and ENF sites**

- Dfc-GRA (1 site)

```{r echo = FALSE, message = FALSE, warning = FALSE,fig.height=10,fig.width=10}
#IT-Tor
df_andPlot_Dfc_GRA<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","GRA","IT-Tor",c(2009:2014))
length(df_andPlot_Dfc_GRA$p_isevent)
p_diffYears<-df_andPlot_Dfc_GRA$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],p_diffYears[[3]],
          p_diffYears[[4]],p_diffYears[[5]],p_diffYears[[6]],        
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)
```

- Dfc-ENF (6 sites)
```{r echo = FALSE, message = FALSE, warning = FALSE,fig.height=10,fig.width=10}
#CA-Man
df_andPlot_Dfc_ENF1<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","CA-Man",c(2000:2003,2007:2008))
length(df_andPlot_Dfc_ENF1$p_isevent)
p_diffYears<-df_andPlot_Dfc_ENF1$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],p_diffYears[[3]],
          p_diffYears[[4]],p_diffYears[[5]],p_diffYears[[6]],
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)

#CA-NS2
df_andPlot_Dfc_ENF2<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","CA-NS2",c(2002,2004))
length(df_andPlot_Dfc_ENF2$p_isevent)
p_diffYears<-df_andPlot_Dfc_ENF2$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)

#CA-NS4
df_andPlot_Dfc_ENF3<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","CA-NS4",c(2005))
length(df_andPlot_Dfc_ENF3$p_isevent)
p_diffYears<-df_andPlot_Dfc_ENF3$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)

#CA-NS5
df_andPlot_Dfc_ENF4<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","CA-NS5",c(2002,2004:2005))
length(df_andPlot_Dfc_ENF4$p_isevent)
p_diffYears<-df_andPlot_Dfc_ENF4$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],p_diffYears[[3]],
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)

#CA-Qfo
df_andPlot_Dfc_ENF5<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","CA-Qfo",c(2004:2010))
length(df_andPlot_Dfc_ENF5$p_isevent)
p_diffYears<-df_andPlot_Dfc_ENF5$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],p_diffYears[[3]],
          p_diffYears[[4]],p_diffYears[[5]],p_diffYears[[6]],
          p_diffYears[[7]],
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)

#FI-Hyy
df_andPlot_Dfc_ENF6<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","FI-Hyy",c(2000:2014))
length(df_andPlot_Dfc_ENF6$p_isevent)
p_diffYears<-df_andPlot_Dfc_ENF6$p_isevent
p_merge<-plot_grid(p_diffYears[[1]],p_diffYears[[2]],p_diffYears[[3]],
          p_diffYears[[4]],p_diffYears[[5]],p_diffYears[[6]],
          p_diffYears[[7]],p_diffYears[[8]],p_diffYears[[9]],
          p_diffYears[[10]],p_diffYears[[11]],p_diffYears[[12]],
          p_diffYears[[13]],p_diffYears[[14]],p_diffYears[[15]],
          labels = "auto",ncol=3,label_size = 12,align = "hv")
plot(p_merge)

```

### step3: save the data that label with "is_event"
```{r echo = FALSE, message = FALSE, warning = FALSE}
# ##(1)Cfb-DE-Hai
# df_andPlot_Cfb<-sep_data_indiffY_sepMismatch(ddf,0.05,"Cfb","DBF","DE-Hai",c(2000:2012))
# 
# ##(2)Dfb
# #Dfb_MF_US-Syv
# df_andPlot_Dfb_MF<-sep_data_indiffY_sepMismatch(ddf,0.05,"Cfb","DBF","US-Syv",c(2002:2006,2013:2014))
# #dfb_DBF
# df_andPlot_Dfb_DBF1<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfb","DBF","US-UMB",c(2000:2014))
# df_andPlot_Dfb_DBF2<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfb","DBF","US-UMd",c(2008:2014))
# df_andPlot_Dfb_DBF3<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfb","DBF","US-WCr",c(2000:2006,2011:2014))
# # df_andPlot_Dfb_DBF4<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfb","DBF","US-Wi3",c(2002,2004))  # no years
# 
# ##(3)Dfc
# #Dfc_GRA_IT-Tor
# df_andPlot_Dfc_GRA<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","GRA","IT-Tor",c(2009:2014))
# #Dfc_ENF
# df_andPlot_Dfc_ENF1<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","CA-Man",c(2000:2003,2007:2008))
# df_andPlot_Dfc_ENF2<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","CA-NS2",c(2002,2004))
# df_andPlot_Dfc_ENF3<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","CA-NS4",c(2005))
# df_andPlot_Dfc_ENF4<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","CA-NS5",c(2002,2004:2005))
# df_andPlot_Dfc_ENF5<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","CA-Qfo",c(2004:2010))
# df_andPlot_Dfc_ENF6<-sep_data_indiffY_sepMismatch(ddf,0.05,"Dfc","ENF","FI-Hyy",c(2000:2014))
#put the data together
#do not add Dfb_DBF4
df_all<-list(df_andPlot_Cfb,
    df_andPlot_Dfb_MF,
    df_andPlot_Dfb_DBF1,df_andPlot_Dfb_DBF2,df_andPlot_Dfb_DBF3,
    df_andPlot_Dfc_GRA,
    df_andPlot_Dfc_ENF1,df_andPlot_Dfc_ENF2,df_andPlot_Dfc_ENF3,df_andPlot_Dfc_ENF4,df_andPlot_Dfc_ENF5,df_andPlot_Dfc_ENF6)

names(df_all)<-c("DE-Hai","US-Syv","US-UMB","US-UMd","US-WCr","IT-Tor","CA-Man","CA-NS2","CA-NS4","CA-NS5","CA-Qfo","FI-Hyy")
ddf_labeled<-c()
for(i in 1:length(df_all)){
  ddf_labeled<-rbind(ddf_labeled,df_all[[i]]$df_agg)
}
#save the data:
save(ddf_labeled,file=paste0(path,"ddf_labeled_norm_kls_newmethod.RDA"))
save(df_all,file=paste0(path,"df_all_results/df_all_norm_kls_newmethod.RDA"))

```

## Summary

### steps to determine the "is_event" period

    
    **Step1: normlization for all the years in one site**
     
    #normalized the gpp_obs and gpp_mod using the gpp_max(95 percentile of gpp)
    
    
    **Step 2:Determine the green-up period for each year(using spline smoothed values):** 
    
    #followed analysis is based on the normlized "GPP_mod"time series(determine earlier sos)
    
    - using the normalized GPP_mod to determine sos,eos and peak of the time series (first to use the klosterman et al., method. in some years, if the klosterman method dose not work, then to use the threshold (percentile 10 of amplitude) to determine the sos and eos in this study. We selected the GPP_mod to determine the phenophases as genearlly we can get earlier sos compared to GPP_obs--> we can have larger analysis period
    
    
    **Step 3:rolling mean of GPPobs and GPPmod for data for all the years(moving windown:5,7,10, 15, 20days)**
    
    **also for the data beyond green-up period--> the code of this steps moves to second step**
    
    - at the end, I select the 20 days windows for the rolling mean
    
    
    **Step 4:Fit the Guassian norm distribution for residuals beyond the green-up period**
    
    - The reason to conduct this are: we assume in general the P-model assume the GPP well outside the green-up period (compared to the observation data). 
    
    - But in practise, the model performance is not always good beyond the green-up period-->I tested three data range:
    
      a. [peak,265/366]
      
      b. DoY[1, sos]& DOY[peak,365/366]
      
      c. [1,sos] & [eos,365/366]
      
    I found the using the data range c, the distrbution of biase (GPP_mod - GPP_obs) is more close to the norm distribution, hence at end of I used the data range c to build the distribution.
    
    
    **step 5:determine the "is_event" within green-up period**
    
      - After some time of consideration, I took following crition to determine the "is_event":
      
        1) during the green-up period (sos,peak)-->the data with GPP biases bigger than 1.2 SD are classified as the "GPP overestimation points"
        
        2) For "GPP overstimation points", thoses are air temparture is less than 10 degrees will be classified as the "is_event". I selected 10 degree as the crition by referring to the paper Duffy et al., 2021 and many papers which demonstrate the temperature response curve normally from 10 degree (for instance: Lin et al., 2012)
      
        References:
        
        Duffy et al., 2021:https://advances.sciencemag.org/content/7/3/eaay1052
        
        Lin et al., 2012:https://academic.oup.com/treephys/article/32/2/219/1657108
      
      
    **step 6:Evaluation "is_event"-->visualization and stats**
    
     - two ways to evaluate if "is_event" is properly determined:
     
      1) visulization
      
      2) stats:
      $$
      Pfalse = \frac{days(real_{(is-event)})}{days(flagged_{(is-event)})} 
      $$